% ============================================================
% main.tex — Submission 281
% "One Flood Substrate, Six Decision Geometries:
%  A Diagnostic Benchmark for Spatial AI [Experiment]"
% SIGSPATIAL 2026, Experiment/Benchmark & Experience track (<=10pp)
%
% OVERLEAF: New Project -> Upload Project -> drag the zip.
%   Compiler: pdfLaTeX. Main document: main.tex.
%   acmart class is built into Overleaf; not vendored here.
%
% BLINDING: default is single-blind (author block visible), matching the
%   submission record. If the track is double-blind, add `anonymous` to the
%   \documentclass options below and the author block auto-hides.
%
% FORMAT: sigconf is the SIGSPATIAL default. Confirm against the 2026
%   Experiment-track CFP; if different, change the class option on the
%   next line only.
% ============================================================
\documentclass[sigconf,nonacm]{acmart}
% ^ `nonacm` removes the ACM copyright/DOI block for clean DRAFT compilation.
%   DO NOT submit with nonacm without checking the SIGSPATIAL 2026 CFP: ACM
%   venues typically want plain [sigconf] (or [sigconf,review]) for the
%   submitted PDF. Confirm the required options before any reviewer-facing
%   build, and do not alter the template to gain space.
%   For double-blind: add `anonymous`.

\usepackage{booktabs}
\usepackage{pifont}      % \ding{51} \ding{55} in the availability matrix
\usepackage{graphicx}
\usepackage{xcolor}

% --- Placeholder + symbol macros (single source of truth) -----------------
% Every UNMEASURED number is wrapped in \PH{} and rendered red so it cannot
% be mistaken for a real result. Delete \PH once all results are locked.
\providecommand{\PH}[1]{\textcolor{red}{#1}}
\providecommand{\Rsup}{S_{\text{sup}}}
\providecommand{\TRF}{\mathrm{TRF}}  % Task Residual Floor (ADR-042, formerly N_ceiling)

% --- Front matter ----------------------------------------------------------
\begin{document}

\title{One Flood Substrate, Six Decision Geometries: A Diagnostic Benchmark for Spatial AI [Experiment]}
% NOTE (camera-ready): the [Experiment] suffix is required by the CFP at
% submission and removed at camera-ready.

\author{Rudolph A. Martin}
\affiliation{%
  \institution{Next Shift Consulting LLC}
  \city{Madison}
  \state{Wisconsin}
  \country{USA}
}
\email{rudy@nextshiftconsulting.com}

% NAMING NOTE: the abstract introduces the benchmark as "GeoRSCT," which is
% also the name of the NeurIPS D&B submission (martin2026georsct). Before
% camera-ready, disclose the relationship explicitly (this paper = GeoRSCT
% instantiated on a flood substrate; cite martin2026georsct as the
% machinery/validation) so the two are not read as duplicate submissions.

\begin{abstract}
Spatial AI benchmarks often reduce flood modeling to one held-out score.
That is insufficient for crisis decision support. A flood dataset may
support prediction, ranking, clustering, transfer, relational propagation,
or allocation, and each task can fail for a different spatial reason.

We introduce GeoRSCT, a diagnostic benchmark for evaluating which
flood-decision tasks a spatial dataset can support. GeoRSCT uses a
multi-event flood substrate spanning multiple metropolitan regions, with
radar and satellite precipitation, terrain, land cover, drainage networks,
levee and urban infrastructure, insurance claims, citizen-reported impacts,
and per-feature provenance and field-status annotations. The benchmark
organizes evaluation around six spatial decision geometries: prediction,
ranking, clustering, transfer, relational propagation, and allocation.

GeoRSCT does not ask only which model scores best. It asks whether the
evidence is sufficient for the decision being made: whether rankings change
under infrastructure context, whether model errors cluster spatially,
whether models transfer across storms and cities, whether drainage or
adjacency assumptions hold across built infrastructure, and whether rainfall
timing gives enough lead time for resource allocation. The result is a
diagnostic account of what a flood dataset supports, what it does not, and
where the evidence is missing or provisional --- surfacing, before a model
is trusted, the points at which a crisis manager would be acting without
sufficient spatial support.
\end{abstract}

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002951.10003227.10003351</concept_id>
<concept_desc>Information systems~Data mining</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10010147.10010178</concept_id>
<concept_desc>Computing methodologies~Machine learning</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}
\ccsdesc[500]{Information systems~Data mining}
\ccsdesc[300]{Computing methodologies~Machine learning}
% TODO: replace with the CCS concepts you select from the ACM CCS tool.

\keywords{Spatial AI, GeoAI, spatial machine learning, spatial data
  infrastructure, geospatial benchmarks, data contracts, reproducible
  benchmarks, flood decision support, evidence sufficiency, disaster response}

\maketitle

% --- Body ------------------------------------------------------------------
\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_problem_formulation}
\input{sections/04_substrate}
\input{sections/05_method}
\input{sections/06_evaluation}   % <-- the one fully-drafted section
\input{sections/07_results}
\input{sections/08_conclusion}

% --- Bibliography ----------------------------------------------------------
% SIGSPATIAL: 10 pages excluding references; up to 2 appendix pages AFTER
% references. Bibliography therefore precedes the appendix.
\bibliographystyle{ACM-Reference-Format}
\bibliography{refs}

% --- Appendix (after references; <=2 pages) --------------------------------
\appendix
\input{appendix}
\input{sections/appendix_doe}
\input{sections/appendix_notation}

\end{document}
